### 2020-06-15 notes/notes.tex
June 15 notes

%        File: notes.tex
%     Created: Mon Jun 15 09:00 AM 2020 E
%

\documentclass[english]{article}
\usepackage[latin9]{inputenc}
\usepackage[letterpaper]{geometry}
\geometry{verbose}
\usepackage{textcomp}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{siunitx}
\usepackage{subfigure}
\usepackage[english]{babel}
\usepackage{braket}
\usepackage{hyperref}
\newcommand{\bibliographytext}{References}

\begin{document}
\title{Research Notes}
\author{Stuart Thomas and Chris Monahan}
\date{Last updated \today}
\maketitle

\section{Monday, June 15}
\begin{enumerate}
    \item Preliminary meeting
    \item We will begin with $\phi^4$ term due to lower energy bound.
    \item Beginning with code in Python, switch to C/C++ if necessary.
\end{enumerate}

\end{document}