### 2021-04-19 final-report.latex/main.tex

\documentclass[12pt]{report}
\usepackage{amsmath}
\usepackage{tabularx}
\usepackage{amsfonts}
\usepackage{hyperref}
\usepackage{tikz}
\usetikzlibrary{arrows.meta}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{listings}
\setlength{\oddsidemargin}{0.5in}
\setlength{\evensidemargin}{0.5in}
\setlength{\textwidth}{6.0in}
\setlength{\topmargin}{.75in}
\setlength{\textheight}{8.6in}

\newcommand{\e}{\vec e}
\newcommand{\citeneeded}{\footnote{\textit{citation needed}}}
\newcommand{\half}{\frac{1}{2}}
\newcommand{\codeexample}[1]{
    \section{\texttt{#1}}
    \lstinputlisting[language=C++]{../code/cpp_code/#1}
}

% Modified by S. Aubin for the class of 2019 Physics theses

%
% This is a LaTeX template for an undergraduate honors or
%   senior thesis in physics at William & Mary. Created from
%   several theses from previous students of D.S. Armstrong,
%   pared down to bare bones.
%
%   you will need this present file (thesis_template.tex)
%   as well as the file  figure1.eps  to make the whole thing
%   (the latter is just an encapsulated postscript figure)
%     For simplicity, we don't use bibtex here; experts can choose
%   to use that more powerful way of dealing with citations.
%
%   To compile it, do:
% > pdflatex thesis_template.tex
% > pdflatex thesis_template.tex  (yes, you need to do it twice to
%                                  fill in the list of tables, figures
%                                  and references).

\begin{document}
\setlength{\baselineskip}{24pt}
\begin{titlepage}
\LARGE
\begin{center}
    {\bf Topology of the $O(3)$ non-linear sigma model under the gradient flow}\\[2.3cm]

\normalsize
A thesis submitted in partial fulfillment of the requirement \\
for the degree of Bachelor of Science with Honors in \\
Physics from the College of William and Mary in Virginia,\\[0.5cm]
by\\[0.5cm]
Stuart Thomas \\[2.5cm]
\end{center}
\normalsize
\begin{flushright}
%\hfill Accepted for Honors \\[.5cm]
\hfill \hrulefill \\
\hfill \hfill Advisor: Prof. Christopher J. Monahan\\[.6cm]
\hfill \hrulefill \\Prof. Todd Averett \\[.6cm]
\hfill \hrulefill \\Prof. Andreas Stathopoulos \\[.6cm]
\end{flushright}
\begin{center}
Williamsburg, Virginia\\
May 2021
\end{center}
\end{titlepage}

\setlength{\topmargin}{0.0in}

\pagenumbering{roman}
\tableofcontents
\chapter*{Acknowledgments}
\addcontentsline{toc}{chapter}{Acknowledgments}
\paragraph
\indent I would like to thank my advisor, Prof. Christopher Monahan. His constant helpfulness has made this research possible and enjoyable. I would also like to thank the entire faculty of the William \& Mary Physics Department. The amount of knowledge and wisdom they have bestowed upon me has been beyond my imagination.

I also acknowledge the help the High Performance Computing group at William \& Mary, whose cluster performed many of these simulations. And finally, I would like to acknowledge the financial help of the 1693 Scholars Program which funded the summer portion of this research.

\listoffigures
\addcontentsline{toc}{chapter}{List of Figures}
\listoftables
\addcontentsline{toc}{chapter}{List of Tables}

\begin{abstract}
\setcounter{page}{5}
\addcontentsline{toc}{chapter}{Abstract}
\paragraph
\indent The $O(3)$ non-linear model features topological effects such as instantons but cannot be perturbatively renormalized. Other nonperturbative theories such as QCD have motivated the development of lattice quantum field theory, where fields are simulated as discrete arrays and extrapolated to the continuum limit. Unfortunately, operators such as the topological susceptibility diverge in this limit. The gradient flow is a strong contender to remove these infinites, having prior success in QCD. We confirm a previous result\cite{bietenholz2018} that the gradient flow does not make the topological susceptibility infinite. We also study the nontrivial field theory by introducing a $\theta$ term into the action and plot the topological charge as a function of $\theta$.

\end{abstract}

\pagenumbering{arabic}
\chapter{Introduction}
\input{intro.tex}
\chapter{Theory}
\input{theory.tex}
\chapter{Methods}
\label{sec:methods}
\input{methods.tex}
\chapter{Results}
\input{results.tex}
\chapter{Conclusion \& Outlook}
\input{conclusion.tex}
\appendix
\chapter{C++ NLSM Monte Carlo Program}
\lstset{language=C++,
                basicstyle=\scriptsize\ttfamily,
                keywordstyle=\color{blue}\ttfamily,
                stringstyle=\color{red}\ttfamily,
                commentstyle=\color{green}\ttfamily,
                morecomment=[l][\color{magenta}]{\#},
                breaklines=true,
                postbreak=\mbox{\textcolor{red}{$\hookrightarrow$}\space}
}

\codeexample{sweep.h}
\codeexample{sweep.cpp}
\codeexample{lattice.h}
\codeexample{lattice.cpp}
\codeexample{phi.h}
\codeexample{phi.cpp}
\codeexample{observables.h}
\codeexample{constants.h}

\bibliographystyle{unsrt}
\bibliography{library}

\end{document}