### 2021-04-22 final-report.latex/main.tex

\documentclass[12pt]{report}
\usepackage{amsmath}
\usepackage{tabularx}
\usepackage{amsfonts}
\usepackage{hyperref}
\usepackage{tikz}
\usetikzlibrary{arrows.meta}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{listings}
%\usepackage[%
    %style=numeric, 
    %sorting=none,
    %doi=false,
    %url=false,
    %giveninits=true,
    %eprint=false,
    %isbn=false,
    %clearlang=true,
    %date=year,
    %hyperref]{biblatex}
\usepackage[style=phys,biblabel=brackets,maxbibnames=3]{biblatex}
%\renewbibmacro{in:}{%
  %\ifentrytype{article}{}{\printtext{\bibstring{in}\intitlepunct}}}

\setlength{\oddsidemargin}{0.5in}
\setlength{\evensidemargin}{0.5in}
\setlength{\textwidth}{6.0in}
\setlength{\topmargin}{.75in}
\setlength{\textheight}{8.6in}

\newcommand{\logchi}{7.4 }
\newcommand{\powchi}{8.8 }

\addbibresource{library.bib}

\newcommand{\e}{\vec e}
\newcommand{\half}{\frac{1}{2}}
\newcommand{\codeexample}[1]{
    \section{\texttt{#1}}
    \lstinputlisting[language=C++]{../code/cpp_code/#1}
}

% Modified by S. Aubin for the class of 2019 Physics theses

%
% This is a LaTeX template for an undergraduate honors or
%   senior thesis in physics at William & Mary. Created from
%   several theses from previous students of D.S. Armstrong,
%   pared down to bare bones.
%
%   you will need this present file (thesis_template.tex)
%   as well as the file  figure1.eps  to make the whole thing
%   (the latter is just an encapsulated postscript figure)
%     For simplicity, we don't use bibtex here; experts can choose
%   to use that more powerful way of dealing with citations.
%
%   To compile it, do:
% > pdflatex thesis_template.tex
% > pdflatex thesis_template.tex  (yes, you need to do it twice to
%                                  fill in the list of tables, figures
%                                  and references).

\begin{document}
\setlength{\baselineskip}{24pt}
\begin{titlepage}
\LARGE
\begin{center}
    {\bf Topology of the $O(3)$ non-linear sigma model under the gradient flow}\\[2.3cm]

\normalsize
A thesis submitted in partial fulfillment of the requirement \\
for the degree of Bachelor of Science with Honors in \\
Physics from the College of William and Mary in Virginia,\\[0.5cm]
by\\[0.5cm]
Stuart Thomas \\[2.5cm]
\end{center}
\normalsize
\begin{flushright}
%\hfill Accepted for Honors \\[.5cm]
\hfill \hrulefill \\
\hfill \hfill Advisor: Prof. Christopher J. Monahan\\[.6cm]
\hfill \hrulefill \\Prof. Todd Averett \\[.6cm]
\hfill \hrulefill \\Prof. Andreas Stathopoulos \\[.6cm]
\end{flushright}
\begin{center}
Williamsburg, Virginia\\
May 2021
\end{center}
\end{titlepage}

\setlength{\topmargin}{0.0in}

\pagenumbering{roman}
\tableofcontents
\chapter*{Acknowledgments}
\addcontentsline{toc}{chapter}{Acknowledgments}
\paragraph
\indent I would like to thank my advisor, Prof. Christopher Monahan. His constant helpfulness has made this research possible and enjoyable. I would also like to thank the entire faculty of the William \& Mary Physics Department. The amount of knowledge and wisdom they have bestowed upon me has been beyond my imagination.

I also acknowledge the help the High Performance Computing group at William \& Mary, whose cluster performed many of these simulations. And finally, I would like to acknowledge the financial help of the 1693 Scholars Program which funded preliminary research over the summer.

\listoffigures
\addcontentsline{toc}{chapter}{List of Figures}
\listoftables
\addcontentsline{toc}{chapter}{List of Tables}

\begin{abstract}
\setcounter{page}{5}
\addcontentsline{toc}{chapter}{Abstract}
\paragraph
\indent Quantum field theory is an extraordinarily successful framework that describes phenomena in particle physics and condensed matter. The $O(3)$ non-linear sigma model (NLSM) is a specific theory used in both of these fields, describing ferromagnets and acting as a prototype for the strong nuclear force. It features topologically stable configurations known as instantons which cannot continuously evolve to the ground state. The topological susceptibility is a parameter that describes this stability and is predicted to vanish in physical theories, however numerical simulations find that the topological susceptibility diverges in the continuum limit \cite{berg1981}. This issue has motivated the application of the ``gradient flow'', a smoothing of high-frequency modes. We study the effect of the gradient flow on this divergence, finding that it persists as a logarithmic divergence. This result supports a previous study \cite{bietenholz2018} and indicates that either the definition of topological charge is problematic or the NLSM has no well-defined continuum limit. We also study the nontrivial field theory by introducing a $\theta$-term into the action and analyze the topological charge as a function of $\theta$ under the gradient flow.

\end{abstract}

\pagenumbering{arabic}

\chapter{Introduction}
\input{intro.tex}
\chapter{Theory}
\input{theory.tex}
\chapter{Methods}
\label{sec:methods}
\input{methods.tex}
\chapter{Results}
\input{results.tex}
\chapter{Conclusion \& Outlook}
\input{conclusion.tex}
\appendix
\chapter{C++ NLSM Monte Carlo Program}
\lstset{language=C++,
                basicstyle=\scriptsize\ttfamily,
                keywordstyle=\color{blue}\ttfamily,
                stringstyle=\color{red}\ttfamily,
                commentstyle=\color{green}\ttfamily,
                morecomment=[l][\color{magenta}]{\#},
                breaklines=true,
                postbreak=\mbox{\textcolor{red}{$\hookrightarrow$}\space}
}

\codeexample{sweep.h}
\codeexample{sweep.cpp}
\codeexample{lattice.h}
\codeexample{lattice.cpp}
\codeexample{phi.h}
\codeexample{phi.cpp}
\codeexample{observables.h}
\codeexample{constants.h}

%\bibliographystyle{unsrt}
%\bibliography{library}

\printbibliography

\end{document}