### 2021-10-01 proceedings/skeleton.tex
add draft  of proceedings

% Please make sure you insert your
% data according to the instructions in PoSauthmanual.pdf
\documentclass[a4paper,11pt]{article}
\usepackage{pos}

\title{Topology of the $O(3)$ non-linear sigma model under the gradient flow}
%% \ShortTitle{Short Title for header}

\author*[a]{Stuart Thomas}
\author[a]{Christopher Monahan}

\affiliation[a]{Department of Physics, William \& Mary,  Williamsburg, Virginia 23187, USA}

%\affiliation[b]{Department, University,\\
%Street number, City, Country}

\emailAdd{snthomas01@email.wm.edu}
\emailAdd{cjmonahan@wm.edu}

\abstract{
The $O(3)$ non-linear sigma model (NLSM) is a prototypical field theory for QCD and ferromagnetism, featuring topological qualities. Though the topological susceptibility $\chi_t$ should vanish in physical theories, lattice simulations of the NLSM find that $\chi_t$ diverges in the continuum limit. We study the effect of the gradient flow on this quantity using a Markov Chain Monte Carlo method, finding that a logarithmic divergence persists. This result supports a previous study and indicates that either the definition of topological charge is problematic or the NLSM has no well-defined continuum limit. We also introduce a $\theta$-term and analyze the topological charge as a function of $\theta$ under the gradient flow.}

\FullConference{%
 The 38th International Symposium on Lattice Field Theory, LATTICE2021
  26th-30th July, 2021
  Zoom/Gather@Massachusetts Institute of Technology
}

\begin{document}
\maketitle

\section{...}

\begin{thebibliography}{99}
\bibitem{...}
....

\end{thebibliography}

\end{document}